---
title: "NOAA Ecosystem Recovery"
author: "Ian Combs"
date: "1/11/2021"
output:
  html_document:
    theme: flatly
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'

---


#### version: `r library(magrittr)` `r Sys.Date() %>% format(format="%B %d, %Y")`

#### [GitHub repository](https://github.com/icombs2017/photogrammetryNOAA.git){target="_blank"}

# About this document
***
#### All analyses preformed with R version `r getRversion()`.


# Basic setup of R environment
***

```{r, setup, include = FALSE}
#setup rmarkdown environment first
knitr::opts_chunk$set(warning = FALSE, fig.align = 'left')
knitr::opts_knit$set(root.dir = '../data')
options(width = 88)
```

## Loading required packages
For the following analyses we will require the use of a number of different R packages. We can use the following code to quickly load in the packages and install any packages not previously installed in the R console.

```{r, load packages, include = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
setwd("../data")

if (!require("pacman")) install.packages("pacman")
pacman::p_load("cowplot", "flextable", "ggrepel", "ggspatial", "lubridate", "officer", "paletteer", "patchwork", "rgdal", "rnaturalearth", "sf", "tidyverse", "gdata", "rgeos", "readxl", "ggstar", "cowplot", "ggpubr", "googledrive")
pacman::p_load_gh("eliocamp/ggnewscale")
pacman::p_load_gh("ropensci/rnaturalearthhires")
```
<br>

# Map of study sites

<br>

```{r, map data}

drive_download(
  as_id("1aXcmLPSRHAlh7N6XhM0g9GKJb4JtOpdJL35dL3Vd678"),
  path = "../data/metadata/siteCoordinates.xlsx", 
  overwrite = TRUE)



sites = read_excel("../data/metadata/siteCoordinates.xlsx", sheet = 2)
sites$site <- as.factor(sites$Site)
sites$species <- as.factor(sites$Species)
levels(sites$site)
sites$`Date Collected` <- as.Date(sites$`Date Collected` )
sites <- sites %>% 
  filter(`Date Collected` > "2022-06-30")



# Now I'm making sure I'm just including the EDR sites
sites <- sites %>% filter(Project == "NFWF")
drop.levels(sites)

sites$reef <- sites$`Iconic Reef?`    



#this is a cheap way of doing this, but i just needed a single coordinate for the site as a whole, so subsetting one site from EDR and using that for my "EDR" designation
edr <- sites %>% filter(site == "T_11")



fknmsBounds <- read_sf("../data/shp/FKNMS_Boundary/FKNMS_Boundary.shp") %>% st_transform(crs = 4326)
flKeys <- read_sf("../data/shp/flKeys.shp") %>% st_transform(crs = 4326)
states = st_as_sf(ne_states(country = c("United States of America")), scale = "large")
countries = st_as_sf(ne_countries(country = c("Cuba", "Mexico", "The Bahamas"), scale = "large"))
florida = read_sf("../data/shp/flKeys.shp") %>% st_transform(crs = 4326)
bathy = read_sf("../data/shp/flBathy.shp") %>% st_transform(crs = 4326) %>% subset(subset = DATASET %in% c("fl_shelf", "fl_coast"))
tortugasBathy = read_sf("../data/shp/tortugasBathy.shp") %>% st_transform(crs = 4326)

```
<br>

Next we build a hi-res polygon of FL with the study site marked and a zoomed in map of the colony locations. We use `ggspatial` to add a north arrow and scale bar to the main map.
```{r, maps}
flPal = paletteer_d("vapoRwave::jazzCup")[c(2:5)]
boundPal = c("gray30", paletteer_d("vapoRwave::vapoRwave")[10])

floridaMap = ggplot() +
  geom_sf(data = states, fill = "white", size = 0.25) +
  geom_sf(data = countries, fill = "white", size = 0.25) +
  geom_sf(data = fknmsBounds, alpha = 0.1) +
  geom_rect(aes(xmin = -82.216, xmax = -81.3464, ymin = 24.2806, ymax =24.8452), color = paletteer_d("vapoRwave::vapoRwave")[6], fill = NA, size = 1) +
  coord_sf(xlim = c(-89, -77), ylim = c(22, 31)) +
  theme_bw() +
  theme(legend.title = element_text(size = 9, face = "bold"),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_rect(fill = "aliceblue"),
        panel.border = element_rect(color = "black", size = 0.75, fill = NA),
        legend.text = element_text(size = 9),
        legend.position = "none",
        plot.background = element_blank())

mainKeys <- ggplot() +
  geom_sf(data = flKeys$geometry, fill = "white", color = "gray30") +
    geom_sf(data = fknmsBounds, fill = "black", alpha = 0.1) +
  geom_sf(data = bathy$geometry, color = "gray75", size = 0.25) +
  coord_sf(xlim = c(-81.955348, -81.689479), ylim = c(24.448239,  24.616647)) +
  geom_star(data = edr, aes(x = long_x, y = lat_y, starshape = reef, fill = reef), size = 10)+
  #Left Segment
  geom_segment(aes(x = -81.86, y = 24.47, xend = -81.900, yend = 24.485), size = 0.25) +
  # Right Segment
  geom_segment(aes(x = -81.825, y = 24.47, xend = -81.781, yend = 24.485), size = 0.25) +
  geom_rect(aes(xmin = -81.86, xmax = -81.825, ymin =24.449, ymax =24.47), color = paletteer_d("vapoRwave::vapoRwave")[6], fill = NA, size = 1) +
  annotation_scale(location = "bl") +
  annotation_north_arrow(location = "br", which_north = "true", style = north_arrow_minimal()) +
  guides(fill = guide_legend(override.aes = list(shape = 22, color = NA, size = 4), ncol = 2, order = 1), shape = guide_legend(override.aes = list(size = 3), order = 2), color = guide_legend(override.aes = list(fill = "black", alpha = 0.1), order = 3)) +
  scale_x_continuous(breaks = c(seq(-82, -80, by = .05))) +
  scale_y_continuous(breaks = c(seq(24, 25, by = .05))) +
theme_bw() +
  theme(panel.background = element_rect(fill = "aliceblue"),
        panel.border = element_rect(color = "black", size = 0.75, fill = NA),
        plot.background = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_line(color = "black"),
        axis.text = element_text(color = "black"),
        legend.position = "none")
  

edrSites <- ggplot() +
  geom_sf(data = flKeys$geometry, fill = "white", color = "gray30") +
    geom_sf(data = fknmsBounds, fill = "black", alpha = 0.2) +
  geom_sf(data = bathy$geometry, color = "grey30", size = 0.25) +
  geom_point(data = sites, aes(x = long_x, y = lat_y, fill = site, shape = species), size = 2, alpha = 1) +
  scale_shape_manual(values = c(21,22), name = "Species", labels = c("ACER", "APAL")) +
  guides(fill = guide_legend(override.aes = list(shape = 22, color = NA, size = 4), ncol = 2, order = 1), shape = guide_legend(override.aes = list(size = 3), order = 2), color = guide_legend(override.aes = list(fill = "black", alpha = 0.1), order = 3)) +
  coord_sf(xlim = c(-81.8459, -81.8427), ylim = c(24.45856,  24.4601)) +
  annotation_scale(location = "bl") +
  scale_fill_discrete(name = "Site")+
  theme_bw() +
  theme(legend.title = element_text(size = 9, face = "bold"),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_rect(fill = "aliceblue"),
        panel.border = element_rect(color = "black", size = 0.75, fill = NA),
        legend.position = "none",
        legend.direction = "vertical",
        legend.box = "horizontal",
        plot.background = element_blank())

# edr_legend <- cowplot::get_legend(edrSites)
# edr_legend <- ggpubr::as_ggplot(edr_legend)


#In order for this to work... you have to change legend.position = "none" to legend.position = "right" (or wve direction you want) the legend from EDR sites above... working to make this better
leg <- get_legend(edrSites)
edrLeg <- as_ggplot(leg)

edrMap = ggdraw() +
  draw_plot(mainKeys) +
  draw_plot(floridaMap, x = .08, y = 0.6, width = 0.25, height = 0.25) +
  draw_plot(edrSites, x =.28, y = .25, width = 0.4, height = 0.4)

# edrNew <- edrMap / edrLeg + plot_layout(heights = unit(c(5,1), c('cm')))

ggsave("../figures/figure1Jitter.png", plot = edrMap, width = 16, height = 16, units = "cm", dpi = 600)

ggsave("../figures/figure1.png", plot = edrMap, width = 16, height = 16, units = "cm", dpi = 600)

ggsave("../figures/figure1Legend.png", plot = edrLeg, width = 16, height = 16, units = "cm", dpi = 600)


```





<br>


```{r, siteTable, include = TRUE}

friedTab = data.frame("Test" = "Total Colony Size", "Comparison" = " ", "Statistic" = fried.total[["statistic"]][["Friedman chi-squared"]], "p.value" = fried.total$p.value[1]) %>% 
  add_row("Test" = " ", "Comparison" = c("T1 - T2", "T1 - T3", "T1 - T4", "T2 - T3", "T2 - T4", "T3 - T4"), "Statistic" = post.total.statistic.melt$value, "p.value" = p.total.melt$value) %>% 
  add_row("Test" = "Healthy Tissue Area", "Comparison" = " ", "Statistic" =  fried.healthy[["statistic"]][["Friedman chi-squared"]], "p.value" = fried.healthy$p.value[1]) %>% 
  add_row("Test" = " ", "Comparison" = c("T1 - T2", "T1 - T3", "T1 - T4", "T2 - T3", "T2 - T4", "T3 - T4"), "Statistic" = post.healthy.statistic.melt$value, "p.value" = p.healthy.melt$value) %>% 
  add_row("Comparison" = " ", "Test" = "Disease Lesion Area", "Statistic" =  fried.disease[["statistic"]][["Friedman chi-squared"]], "p.value" = fried.disease$p.value[1]) %>% 
  add_row("Test" = " ", "Comparison" = c("T1 - T2", "T1 - T3", "T1 - T4", "T2 - T3", "T2 - T4", "T3 - T4"), "Statistic" = post.disease.statistic.melt$value, "p.value" = p.disease.melt$value) %>% 

mutate_if(is.numeric, round, digits = 3) %>%     
mutate(p.value = replace(p.value, p.value >= 0.05, "ns")) %>%
mutate(p.value = replace(p.value, p.value < 0.001, "< 0.001")) %>% 
mutate_if(is.character, str_replace_all, pattern = "-", replacement = "–") %>%
flextable() %>%
set_header_labels(Data.set = "Data set") %>% 
flextable::compose(part = "header", j = "Statistic", value = as_paragraph("Test Statistic")) %>%
flextable::compose(part = "header", j = "p.value", value = as_paragraph(as_i("p"), "-value")) %>% 
autofit() %>%
font(fontname = "Times New Roman", part = "all") %>%
fontsize(size = 12, part = "all") %>%
bold(part = "header") %>% 
colformat_num(j = "Statistic", digits = 2) %>%
colformat_num(j = "p.value", digits = 4, na_str = "ns") %>% 
align_nottext_col(align = "center", header = TRUE, footer = TRUE) %>% 
align(align = "center", j = "p.value")



siteTable <- sites %>% 
  select(Site, Species, `Area Surveyed (m^2)`, `Date Collected`, TimePoint, lat_y, long_x) %>%
  filter(between(`Date Collected`, as.Date('2022-01-01'), as.Date('2022-06-30'))) %>%
  arrange(`Date Collected`) %>% 
  flextable() %>% 
  set_header_labels(Data.set = "Data set") %>% 
  flextable::compose(part = "header", j = "lat_y", value = as_paragraph("Latitude")) %>% 
  flextable::compose(part = "header", j = "long_x", value = as_paragraph("Longitude")) %>% 
  flextable::compose(part = "header", j = "TimePoint", value = as_paragraph("Time Point")) %>% 
  autofit() %>%
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  flextable::fontsize(size = 12, part = "all") %>%
  bold(part = "header") %>% 
  align_nottext_col(align = "center", header = TRUE, footer = TRUE)






siteDoc = read_docx()
siteDoc = body_add_flextable(siteDoc, value = siteTable)
print(siteDoc, target = "../tables/Table2.docx")
siteTable













```





