---
title: "NFWF VPI"
author: "Ian Combs -- icombs@mote.org"
output:
  html_document:
    theme: flatly
    code_folding: show
    toc: yes
    toc_depth: 3
    toc_float: yes
  pDf_doctument:
      toc: yes
      toc_depth: 3
---

```{r, setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, fig.align = 'left')
library(magrittr)
```

### version: `r Sys.Date() %>% format(format="%B %d, %Y")`

<!-- this is where the DOI would go  [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3675991.svg)](https://doi.org/10.5281/zenodo.3675991)
-->


#### [GitHub repository](https://github.com/icombs2017/photogrammetryNOAA.git){target="_blank"}

###
+++
This is the analysis pipeline for data generated from TagLab annotations of our EDR outplant sites as part of the NWFW project.


+++
### All analyses performed with R verion `r getRversion()`
# Basic setup of R environment
+++
## Loading required packages

For the following analyses we will require the use of a number of different R packages. Most of which can be sourced from CRAN, but some must be downloaded from GitHub. We can use the following code to load in the packages and install any packages not previously installed in the R console. 

```{r, load packages, include = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("ggplot2","officer","ggpubr", "rcompanion", "RColorBrewer", "magrittr","reshape2", "stringr", "plyr", "dplyr", "flextable", "tidyr", "tibble", "vegan", "forcats", "readr", "metafolio", "FSA","viridis", "paletteer", "purrr", "data.table", "ecodist", "glue", "gtools", "patchwork")
pacman::p_load_gh("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
```

<br><br>

# Loading Data
+++

Data is exported from Viscore's VPI scriplet as a .csv file and named according to the file. 
Since Viscore puts the "percent" column as a percentage sign (++%++) R cannot read it in  we are using `dplyr`'s `rename` function to change it to something R can read and that is easier to call and read when we are handling the data. Before we manipulate the data I am first making a vector called ++myColors++ using the function `gg_color_hue()` from the package `metafolio` that will save a vector of colors to the number of factors I have, this is so in our later plots, all the color palettes will be the same. I am also adding ++site++ and ++time++ columns to each subset, combining all subsets into one large dataframe using `bind_rows()`, adding a ++treatment++ column and using the `gsub()` to rename some 
```{r, load data, include = TRUE, results = 'hide'}


read_plus <- function(flnm) {
    read_csv(flnm) %>% 
        mutate(filename = flnm)
}

vpiData1 <-
    list.files(path = "../data/vpiData", pattern = "+.csv", 
               full.names = T) %>% 
    map_dfr(~read_plus(.))


#Sort out data to have proper columns for all the info i want, this might need to be changed to figure out the Marks vs Plot thing

vpiData2 <- vpiData1 %>% 
      tidyr::separate(filename, into=c("Date", "Site"),extra = 'merge',sep = "(?=T)") %>%
      tidyr::separate(Date, into=c("remove", "Date"),sep = "../data/vpiData/") %>% 
      tidyr::separate(Site, into = c("Site", "delete"),sep = ".pqs") %>% 
      dplyr::select(-c(remove, delete)) %>% 
      mutate_if(is.character, str_replace_all, pattern = "_", replacement = "-") %>%
      tidyr::separate(Date, into = c("Date", "remove"), sep = 10) %>% 
      dplyr::select(-remove) %>% 
      tidyr::separate(Site, into = c("Site", "Plot"), sep = "(?=plot)", extra = "merge") %>% 
      tidyr::separate(Site, into = c("Site", "Mark"), sep = "(?=mark)", extra = "merge") %>% 
      tidyr::separate(Plot, into = c("Plot", "Mark"), sep = "(?=mark)", extra = "merge" ) %>%
      mutate(species = if_else(grepl("AP", Site), "APAL", "ACER")) %>% 
      dplyr::rename(abundance = `%`) %>%
      tidyr::drop_na(abundance) %>% 
      arrange(desc(abundance))




#Changing the Date to factors
vpiData2$Date<-as.Date(vpiData2$Date)



#Making some corrections in the naming
vpiData3 <- vpiData2 %>%
  mutate_all(~gsub("Crustose coralline algae natans", "Crustose coralline algae", .)) %>%
  mutate_all(~gsub("Cordata", "Chordata", .)) %>%
  mutate_all(~gsub("Rubble", "Bare Substrate", .)) %>%
  mutate(Site = gsub("T-AP-7pt2", "T-AP-7", Site),
         Site = gsub("T-AP-7pt1", "T-AP-7", Site),
         Site = gsub("T-AP-17pt2", "T-AP-17", Site),
         Site = gsub("T-AP-17pt1", "T-AP-17", Site),
         Site = gsub("T-AP-16pt2", "T-AP-16", Site),
         Site = gsub("T-16pt1", "T-16", Site),
         Site = gsub("T-16pt2", "T-16", Site),
         Site = gsub("T-17pt1", "T-17", Site),
         Site = gsub("T-17pt2", "T-17", Site),
         Site = gsub("T-17pt3", "T-17", Site),
         Site = gsub("T-17pt4", "T-17", Site),
         Site = gsub("T-AP-13pt2", "T-AP-13", Site),
         Site = gsub("T-AP-13pt1", "T-AP-13", Site),
         Site = gsub("T-AP-14pt1", "T-AP-14", Site),
         Site = gsub("T-AP-14pt2", "T-AP-14", Site),
         Site = gsub("T-12a", "T-12", Site),
         Site = gsub("T-12b", "T-12", Site),
         Site = gsub("T-13a", "T-13", Site),
         Site = gsub("T-13b", "T-13", Site),
         Site = gsub("T-AP-3a", "T-AP-3", Site),
         Site = gsub("T-18pt1", "T-18", Site),
         Site = gsub("T-18pt2", "T-18", Site)
) %>%
  mutate_all(~gsub("Other|Review|None", NA, .))



# Changing other columns to factors

vpiData4 <- vpiData3 %>%
  mutate(across(c(Site, class, timePoint, species), as.factor)) %>%
  mutate(Date = as.Date(Date)) %>%
  mutate(count = as.numeric(count)) %>% 
  select(Date, Site, timePoint, species, class, count)


# Changing the factor level of timePoint to be chronological
vpiData <- vpiData4 %>%
  mutate(timePoint = factor(timePoint, levels = levels(timePoint)[c(2,1)]))
vpiData



```


<br><br>
# Combining Data
+++
Multiple VPI surveys were conducted in each site, so we are combining all of the sites and creating a new abundance column from that
```{r}

vpiData2 <- vpiData %>% 
  dplyr::filter(class != c("NA")) %>%
  droplevels() %>%
  group_by(Site,timePoint, species, class) %>% 
  dplyr::summarise(totalCount = sum(count)) %>%
  ungroup() %>% 
  group_by(Site,timePoint) %>% 
  dplyr::mutate(totalObservations = sum(totalCount, na.rm = TRUE)) %>%
  ungroup() %>% 
  group_by(Site,timePoint, class) %>% 
  dplyr::mutate(abundance = ((totalCount/totalObservations)+100))


vpiData2 <- data.table(vpiData2)
vpiData2[, `:=`(class, reorder(class, abundance))]
# vpiData2$class <- drop_na(vpiData2$class)



```


<br><br>
# Plotting the data
+++

To visualize the data and what our species composition looks like at our sites, we are going to plot the data in a few different ways. First we are going to just plot those with >1% abundance at our sites.



```{r, stackedBarGraphs, include = TRUE}

ACER <- vpiData2 %>% 
  filter(species == "ACER") %>% 
  filter(abundance > 1) %>% 
  droplevels() %>% 
  mutate(Site = factor(Site, levels = rev(mixedsort(levels(Site)))))

APAL <- vpiData2 %>% 
  filter(species == "APAL") %>% 
  filter(abundance > 1) %>% 
  droplevels() %>% 
  mutate(Site = factor(Site, levels = rev(mixedsort(levels(Site)))))




acerStack1 <- ggplot(ACER, aes(x = timePoint, y = abundance))+
                      geom_bar(aes(fill = factor(class)), position = position_stack(reverse = FALSE), stat = 'identity', color = 'black', alpha = 0.8)+
                      ylim(0,100)+
                      labs(y = "Relative Abundance (%)", fill = "Class", title = "Relative Abundance > 1%", x = "Time Point")+
  scale_x_discrete(labels = c("Baseline", "12 Month"))+
                      # scale_y_reverse()+
                      facet_wrap(.~Site, scales = 'free')





acerStack <- acerStack1 + theme(
    # Title
    plot.title = element_text(size = 40, face = "bold"), 
    # X Axis
    axis.text.x = element_text(size = 15, colour = "black", vjust = 0.5, hjust = 0.5, face= "bold"), 
    axis.title.x = element_blank(),
    # Y Axis
    axis.title.y = element_text(size = 40, face = "bold"), 
    axis.text.y = element_text(colour = "black", size = 30, face = "bold"),
    # Axis Lines and Ticks
    axis.line = element_line(colour = "black"), 
    axis.ticks = element_line(color="black"), 
    # Legend
    legend.title = element_text(size = 40, face = "bold"), 
    legend.text = element_text(size = 36, face = "bold", colour = "black"), 
    legend.position = "right",
    # Grid and Background
    panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
    panel.background = element_rect(fill = '#F5F5F5'), 
    # Text
    text = element_text(size=40, color="black"))



acerStack

apalStack1 <- ggplot(APAL, aes(x = timePoint, y = abundance))+
                      geom_bar(aes(fill = factor(class)), position = position_stack(reverse = FALSE), stat = 'identity', color = 'black', alpha = 0.8)+
                      ylim(0,100)+
                      labs(y = "Relative Abundance (%)", fill = "Class", title = "Relative Abundance > 1%", x = "Time Point")+
  scale_x_discrete(labels = c("Baseline", "12 Month"))+
                      # scale_y_reverse()+
                      facet_wrap(.~Site, scales = 'free')





apalStack <- apalStack1 +theme(
    # Title
    plot.title = element_text(size = 40, face = "bold"), 
    # X Axis
    axis.text.x = element_text(size = 15, colour = "black", vjust = 0.5, hjust = 0.5, face= "bold"), 
    axis.title.x = element_blank(),
    # Y Axis
    axis.title.y = element_text(size = 40, face = "bold"), 
    axis.text.y = element_text(colour = "black", size = 30, face = "bold"),
    # Axis Lines and Ticks
    axis.line = element_line(colour = "black"), 
    axis.ticks = element_line(color="black"), 
    # Legend
    legend.title = element_text(size = 40, face = "bold"), 
    legend.text = element_text(size = 36, face = "bold", colour = "black"), 
    legend.position = "right",
    # Grid and Background
    panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
    panel.background = element_rect(fill = '#F5F5F5'), 
    # Text
    text = element_text(size=40, color="black"))
apalStack

# comboPlot <- acerStack + apalStack + plot_layout(guides = 'collect')


# ggsave("../figures/comboPlot.png", plot = comboPlot, width = 20, height = 15, units = 'in', dpi = 600)

ggsave("../figures/acerStackVPI.png", plot = acerStack, width = 20, height = 15, units = 'in', dpi = 600)
ggsave("../figures/apalStackVPI.png", plot = apalStack, width = 20, height = 15, units = 'in', dpi = 600)




```

# Analysis
We are making dissimilarity matrices in order to understand relatedness between outplant sites, events, etc and visualizing them in a few different ways including PCOAs and NMDSs with final statistical analysis being conducted as a Permuational Analysis of Variance (PERMANOVA) using the package `vegan` and the function `adonis2()`.
```{r, dissimilarityMatrix, include = TRUE}


vpiMat <- vpiData2 %>% 
  select(Site, timePoint, species, class, totalCount) %>% 
  group_by(Site, timePoint, species, class) %>%
  # dplyr::summarise(totalCount = sum(count)) %>%
  na.omit(class) %>% 
  tidyr::unite(sample, 1:3) %>% 
  spread(class,totalCount) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames(var = 'sample')

dist_matrix <- as.matrix(bcdist(vpiMat),byrow = TRUE)


dist_tbl <- dist_matrix %>% 
        as_tibble(rownames = 'samples')

samples <- dist_tbl %>% 
  select(samples)


 samples_lookup <- dist_tbl %>% 
   select(samples) %>% 
   separate(col = samples, into = c("site", "timePoint", "species"), sep = "_", remove = FALSE)
 


  

dist_tbl %>% 
  pivot_longer(cols = -samples, names_to = "b", values_to = "distances") %>% 
  pivot_wider(names_from = "b", values_from = "distances")
 
 

dist_matrix <- dist_tbl %>%
  pivot_longer(cols=-samples, names_to="b", values_to="distances") %>%
  inner_join(., samples_lookup, by="samples") %>%
  inner_join(., samples_lookup, by=c("b" = "samples")) %>%
  # filter(day.x %in% days_wanted & day.y %in% days_wanted) %>%
  select(samples, b, distances) %>%
  pivot_wider(names_from="b", values_from="distances") %>%
  select(-samples) %>%
  as.dist()


```

```{r, PCoA, include = TRUE}


pcoa <- cmdscale(dist_matrix, k = 2, eig = TRUE, add = TRUE) # add k = 3, k = 5 to add columns, default is k=2
positions <- pcoa$points
colnames(positions) <- c("pcoa1", 'pcoa2')
# calculating the amount of variation explained by each dimension
percentExplained <- 100 + (pcoa$eig / sum(pcoa$eig))

#rounding and trimming to add to labels
RoundPercentExplained <- format(round(percentExplained[1:2], digits = 1), nsmall = 1, trim = TRUE)

# creating label vector
labs <- c(glue("PCo 1 ({RoundPercentExplained[1]}%)"),
          glue("PCo 2 ({RoundPercentExplained[2]}%)"))




treatmentCentroid <- positions %>% as_tibble(rownames='samples') %>%
  inner_join(., samples_lookup, by = 'samples') %>% 
  group_by(timePoint) %>% 
  dplyr::summarise(axis1 = mean(pcoa1),
            axis2 = mean(pcoa2))




#
pcoaTbl <- positions %>% as_tibble(rownames='samples') %>%
  inner_join(., samples_lookup, by = 'samples') 

pcoaPlot1 <- ggplot() +
  geom_point(data = pcoaTbl, mapping = aes(x = pcoa1, y = pcoa2, color = timePoint, shape = timePoint), size = 3) +
  geom_point(data = treatmentCentroid, mapping = aes(x = axis1, y = axis2, color = timePoint), shape = 15, size = 5) +
  labs(x = labs[1], y = labs[2])


pcoaPlot2 <- ggplot(pcoaTbl, aes(x = pcoa1, y = pcoa2, color = species, fill = species)) +
  stat_ellipse(geom = 'polygon', alpha = 0.2, level = 0.75, show.legend = FALSE) +
  geom_point(aes(x = pcoa1, y = pcoa2, shape = timePoint), alpha = 1, size = 3) +
  # geom_point(data = treatmentCentroid, mapping = aes(x = axis1, y = axis2, color = treatment), shape = 15, size = 5) +
  labs(x = labs[1], y = labs[2])

pcoaPlot1 <- pcoaPlot2 + theme(
    # Title
    plot.title = element_text(size = 40, face = "bold"),
    # X Axis
    axis.text.x = element_text(angle = 45, hjust = 1),  
    axis.title.x = element_text(size = 20, face = "bold"),
    # Y Axis
    axis.title.y = element_text(size = 20, face = "bold"),
    axis.text.y = element_text(colour = "black", size = 20, face = "bold"),
    # Axis Lines and Ticks
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(color="black"),
    # Legend
    legend.title = element_text(size = 20, face = "bold"),
    legend.text = element_text(size = 20, face = "bold", colour = "black"),
    legend.position = "bottom",
    # Grid and Background
    panel.grid.major = element_line(size = 0.3, linetype = 'solid', colour = "black"),
    panel.background = element_rect(fill = 'white'),
    # Text
    text = element_text(size=20, color="black"))


ggsave("../figures/pcoaPlot.png", plot = pcoaPlot1, width = 20, height = 15, units = 'in', dpi = 600)



tibble(pe = cumsum(percentExplained),
       axis = 1:length(percentExplained)) %>%
  ggplot(aes(x = axis, y = pe))+
  geom_line()+
  coord_cartesian(xlim = c(1,10))




```

```{r, NMDS, include = TRUE}

set.seed(3008)
nmds <- metaMDS(dist_matrix)

scores(nmds) %>% 
  as_tibble(rownames='samples') %>% 
  inner_join(., samples_lookup, by = 'samples') %>% 
  ggplot(aes(x = NMDS1, y = NMDS2, color = timePoint))+
  geom_point()
  




```

```{r, ecologicalDistances, include = TRUE}

ecoMat <- as.matrix(vpiMat)
samples_lookup <- vpiMat %>%
  rownames_to_column(var = 'samples') %>% 
   select(samples) %>% 
   separate(col = samples, into = c("site", "timePoint", "species"), sep = "_", remove = FALSE)


set.seed(3008)
dist <- vegdist(ecoMat, method = 'bray')
nmds <- metaMDS(dist)

scores(nmds) %>% 
  as_tibble(rownames = 'samples') %>%
  inner_join(., samples_lookup, by = 'samples') %>% 
  ggplot(aes(x = NMDS1, y = NMDS2, shape = timePoint, color = species))+
  geom_point()







```

```{r, PERMANOVA, include = TRUE, message = FALSE}

set.seed(9999)

vpiMat1 <- vpiMat %>% 
  select(-c(`Bare Substrate`))



 samplesLookupPerm <- vpiMat1 %>% 
   rownames_to_column(var = "samples") %>% 
   select(samples) %>% 
   separate(col = samples, into = c("site", "timePoint", "species"), sep = "_", remove = FALSE) %>% 
   column_to_rownames(var = "samples")
 
vpiDist <- vegdist(vpiMat1, method = 'bray')

distPerm <- adonis2(vpiDist ~ timePoint+species, data = samplesLookupPerm, permutations = 9999, method = 'bray')

distPerm



```




```{r, PERMANOVAtable, include = TRUE, results = hide}

distPermTab <- distPerm %>%
  as_tibble(rownames = "Comparison") %>%
  dplyr::rename("Pseudo.F" = "F", "p.value" = "Pr(>F)") %>% 
  select(Comparison, Df, `Pseudo.F`, `p.value`) %>% 
  mutate_if(is.numeric, round, digits = 3) %>%
  mutate(p.value = replace(p.value, p.value >= 0.05, "ns")) %>%
  mutate(p.value = replace(p.value, p.value < 0.001, "< 0.001")) %>% 
  mutate_if(is.character, str_replace_all, pattern = "-", replacement = "–") %>%
  flextable() %>%
  set_header_labels(Data.set = "Data set") %>% 
  flextable::compose(part = "header", j = "Df", value = as_paragraph("Df")) %>% 
  flextable::compose(part = "header", j = "Pseudo.F", value = as_paragraph("Pseudo-F")) %>%
  flextable::compose(part = "header", j = "p.value", value = as_paragraph(as_i("p"), "-value")) %>% 
  autofit() %>%
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 12, part = "all") %>%
  bold(part = "header") %>%
  set_table_properties(width = 1, align = "center", layout = "autofit") %>%  
  align(align = 'center', part = "footer")




distDoc = read_docx()
distDoc = body_add_flextable(distDoc, value = distPermTab)
print(distDoc, target = "../tables/distTableVpi.docx")
distPermTab









```






```{r}


indexMatrix <- vpiData2 %>% 
  select(Site, timePoint, species, class, totalCount) %>% 
  group_by(Site, timePoint, species, class) %>%
  # dplyr::summarise(totalCount = sum(count)) %>%
  na.omit(class)


richness <- specnumber(vpiMat)  # Count of non-zero species per row
print(richness)


shannon <- diversity(vpiMat, index = "shannon")  # Shannon index (H')
evenness <- shannon / log(richness)  # Pielou's evenness (J')

simpson <- diversity(vpiMat, index = "simpson")
print(simpson)

inv_simpson <- diversity(vpiMat, index = "invsimpson")
print(inv_simpson)



diversityDF <- data.frame(
  Sample = rownames(vpiMat),  # Keep row names
  Richness = richness,
  Shannon = shannon,
  Simpson = simpson,
  InvSimpson = inv_simpson,
  Evenness = evenness,
  row.names = rownames(vpiMat)  # Assign row names back
) %>% 
  separate(col = Sample, into = c("site", "timePoint", "species"), sep = "_", remove = FALSE) %>% 
  mutate(across(c(site, timePoint, species), as.factor))

aov(Richness ~ timePoint+species+site, data = diversityDF)


# Fit the ANOVA model for Richness
richness_aov <- aov(Richness ~ timePoint + species + site, data = diversityDF)

# Summarize the ANOVA table
anova_richness <- anova(richness_aov)
anova_richness

# Fit the ANOVA model for Shannon
shannon_aov <- aov(Shannon ~ timePoint + species + site, data = diversityDF)

# Summarize the ANOVA table
anova_shannon <- anova(shannon_aov)
anova_shannon


# Fit the ANOVA model for Simpson
simpson_aov <- aov(Simpson ~ timePoint + species + site, data = diversityDF)

# Summarize the ANOVA table
anova_simpson <- anova(simpson_aov)
anova_simpson


# Fit the ANOVA model for Evenness
evenness_aov <- aov(Evenness ~ timePoint + species + site, data = diversityDF)

# Summarize the ANOVA table
anova_evenness <- anova(evenness_aov)
anova_evenness

```



```{r}
# Richness ANOVA
richness_aov <- aov(Richness ~ timePoint + species + site, data = diversityDF)
anova_richness <- anova(richness_aov)

# Convert the ANOVA result to tibble
# Assuming anova_richness is your ANOVA result object
richness_tab <- as_tibble(anova_richness) %>%
  as_tibble(rownames = "Comparison") %>%
  dplyr::rename(Pseudo.F = `F value`, p.value = `Pr(>F)`) %>%
  select(Comparison, Df, Pseudo.F, p.value) %>%
  mutate_if(is.numeric, round, digits = 3) %>%
  mutate(p.value = replace(p.value, p.value >= 0.05, "ns")) %>%
  mutate(p.value = replace(p.value, p.value < 0.001, "< 0.001")) %>%
  mutate_if(is.character, str_replace_all, pattern = "-", replacement = "–")

# Manually add factor names to the 'Comparison' column (e.g., "timePoint", "species", "site")
richness_tab$Comparison <- c("Time Point", "Species", "Site", "Residuals")

# Print the updated table
print(richness_tab)

# Create the flextable for Richness
richness_flextable <- richness_tab %>%
  flextable() %>%
  set_header_labels(Data.set = "Data set") %>% 
  flextable::compose(part = "header", j = "Df", value = as_paragraph("Df")) %>% 
  flextable::compose(part = "header", j = "Pseudo.F", value = as_paragraph("Pseudo-F")) %>%
  flextable::compose(part = "header", j = "p.value", value = as_paragraph(as_i("p"), "-value")) %>% 
  autofit() %>%
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 12, part = "all") %>%
  bold(part = "header") %>%
  set_table_properties(width = 1, align = "center", layout = "autofit") %>%  
  align(align = 'center', part = "footer")

# View the flextable for Richness
richness_flextable

richnessDoc = read_docx()
richnessDoc = body_add_flextable(richnessDoc, value = richness_flextable)
print(richnessDoc, target = "../tables/richnessTable.docx")
richness_flextable

```

```{r, shannon, include = TRUE}


# Shannon ANOVA
shannon_aov <- aov(Shannon ~ timePoint + species + site, data = diversityDF)
anova_shannon <- anova(shannon_aov)


# Convert the ANOVA result to tibble
shannon_tab <- as_tibble(anova_shannon) %>%
  as_tibble(rownames = "Comparison") %>%
  dplyr::rename(Pseudo.F = `F value`, p.value = `Pr(>F)`) %>%
  select(Comparison, Df, Pseudo.F, p.value) %>%
  mutate_if(is.numeric, round, digits = 3) %>%
  mutate(p.value = replace(p.value, p.value >= 0.05, "ns")) %>%
  mutate(p.value = replace(p.value, p.value < 0.001, "< 0.001")) %>%
  mutate_if(is.character, str_replace_all, pattern = "-", replacement = "–")

shannon_tab$Comparison <- c("Time Point", "Species", "Site", "Residuals")

# Create the flextable for Shannon
shannon_flextable <- shannon_tab %>%
  flextable() %>%
  set_header_labels(Data.set = "Data set") %>% 
  flextable::compose(part = "header", j = "Df", value = as_paragraph("Df")) %>% 
  flextable::compose(part = "header", j = "Pseudo.F", value = as_paragraph("Pseudo-F")) %>%
  flextable::compose(part = "header", j = "p.value", value = as_paragraph(as_i("p"), "-value")) %>% 
  autofit() %>%
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 12, part = "all") %>%
  bold(part = "header") %>%
  set_table_properties(width = 1, align = "center", layout = "autofit") %>%  
  align(align = 'center', part = "footer")

# View the flextable for Shannon
shannon_flextable


shannonDoc = read_docx()
shannonDoc = body_add_flextable(shannonDoc, value = shannon_flextable)
print(shannonDoc, target = "../tables/ShannonTable.docx")
shannon_flextable


```




```{r, Simpson, include = TRUE}


# Simpson ANOVA
simpson_aov <- aov(Simpson ~ timePoint + species + site, data = diversityDF)
anova_simpson <- anova(simpson_aov)

# Convert the ANOVA result to tibble
simpson_tab <- as_tibble(anova_simpson) %>%
  as_tibble(rownames = "Comparison") %>%
  dplyr::rename(Pseudo.F = `F value`, p.value = `Pr(>F)`) %>%
  select(Comparison, Df, Pseudo.F, p.value) %>%
  mutate_if(is.numeric, round, digits = 3) %>%
  mutate(p.value = replace(p.value, p.value >= 0.05, "ns")) %>%
  mutate(p.value = replace(p.value, p.value < 0.001, "< 0.001")) %>%
  mutate_if(is.character, str_replace_all, pattern = "-", replacement = "–")


simpson_tab$Comparison <- c("Time Point", "Species", "Site", "Residuals")


# Create the flextable for Simpson
simpson_flextable <- simpson_tab %>%
  flextable() %>%
  set_header_labels(Data.set = "Data set") %>% 
  flextable::compose(part = "header", j = "Df", value = as_paragraph("Df")) %>% 
  flextable::compose(part = "header", j = "Pseudo.F", value = as_paragraph("Pseudo-F")) %>%
  flextable::compose(part = "header", j = "p.value", value = as_paragraph(as_i("p"), "-value")) %>% 
  autofit() %>%
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 12, part = "all") %>%
  bold(part = "header") %>%
  set_table_properties(width = 1, align = "center", layout = "autofit") %>%  
  align(align = 'center', part = "footer")

# View the flextable for Simpson
simpson_flextable



simpsonDoc = read_docx()
simpsonDoc = body_add_flextable(simpsonDoc, value = simpson_flextable)
print(simpsonDoc, target = "../tables/simpsonTable.docx")
simpson_flextable
```

```{r, evenness, include = TRUE}

# Evenness ANOVA
evenness_aov <- aov(Evenness ~ timePoint + species + site, data = diversityDF)
anova_evenness <- anova(evenness_aov)

# Convert the ANOVA result to tibble
evenness_tab <- as_tibble(anova_evenness) %>%
  as_tibble(rownames = "Comparison") %>%
  dplyr::rename(Pseudo.F = `F value`, p.value = `Pr(>F)`) %>%
  # select(Comparison, Df, Pseudo.F, p.value) %>%
  mutate_if(is.numeric, round, digits = 3) %>%
  mutate(p.value = replace(p.value, p.value >= 0.05, "ns")) %>%
  mutate(p.value = replace(p.value, p.value < 0.001, "< 0.001")) %>%
  mutate_if(is.character, str_replace_all, pattern = "-", replacement = "–")

evenness_tab$Comparison <- c("Time Point", "Species", "Site", "Residuals")

# Create the flextable for Evenness
evenness_flextable <- evenness_tab %>%
  flextable() %>%
  set_header_labels(Data.set = "Data set") %>% 
  flextable::compose(part = "header", j = "Df", value = as_paragraph("Df")) %>% 
  flextable::compose(part = "header", j = "Pseudo.F", value = as_paragraph("Pseudo-F")) %>%
  flextable::compose(part = "header", j = "p.value", value = as_paragraph(as_i("p"), "-value")) %>% 
  autofit() %>%
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 12, part = "all") %>%
  bold(part = "header") %>%
  set_table_properties(width = 1, align = "center", layout = "autofit") %>%  
  align(align = 'center', part = "footer")

# View the flextable for Evenness
evenness_flextable




evennessDoc = read_docx()
evennessDoc = body_add_flextable(evennessDoc, value = evenness_flextable)
print(evennessDoc, target = "../tables/evennessTable.docx")
evenness_flextable





```













